---
title: "Take Home Exercise 3a: Modelling Geography of Financial Inclusion with Geographically Weighted Methods"
author: "Jeffrey Lee Shao Lin"
date: "Oct 28 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  freeze: true  #will not re-render all items
  message: false #no warning messages go out
---

# 1. Introduction

According to [Wikipedia](https://en.wikipedia.org/wiki/Financial_inclusion), **financial inclusion** is the availability and equality of opportunities to access financial services. It refers to processes by which individuals and businesses can access appropriate, affordable, and timely financial products and services - which include banking, loan, equity, and insurance products. It provides paths to enhance inclusiveness in economic growth by enabling the unbanked population to access the means for savings, investment, and insurance towards improving household income and reducing income inequality.

# 2. The Task

In this take-home exercise, we are required to build an explanatory model to determine factors affecting financial inclusion by using geographical weighted regression methods.

# 3. The Data

For the purpose of this take-home exercise, two data sets shall be used, they are:

-   [FinScope Uganda 2023](https://fsduganda.or.ug/data-sets-finscope-uganda-2023-survey-report/)

-   The district level boundary GIS data can be downloaded from [geoBoundaries](https://www.geoboundaries.org/) portal

# 4. Importing Packages

Before we start the exercise, we will need to import necessary R packages first. We will use the following packages:

-   [**olsrr**](https://olsrr.rsquaredacademy.com/) package for building OLS and performing diagnostics tests

-   [**GWmodel**](https://cran.r-project.org/web/packages/GWmodel/index.html) package for calibrating geographical weighted family of models

-   [**corrplot**](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html) package for multivariate data visualisation and analysis

-   [sf](https://r-spatial.github.io/sf/index.html) package provides functions to manage, processing, and manipulate **Simple Features**, a formal geospatial data standard that specifies a storage and access model of spatial geometries such as points, lines, and polygons.

-   [**tmap**](https://cran.r-project.org/web/packages/tmap/index.html) which provides functions for plotting cartographic quality static point patterns maps or interactive maps by using [leaflet](https://leafletjs.com/) API.

    Use the code chunk below to install and launch the below R packages.

```{r}
pacman::p_load(olsrr, corrplot, ggpubr, sf, spdep, GWmodel, tmap, tidyverse, gtsummary)
```

# 5. Getting the Data Into R Environment

## 5.1 Importing geospatial data

The geospatial data used in this hands-on exercise is called geoBoundaries-UGA-ADM2. It is in ESRI shapefile format. The shapefile consists of Uganda district level boundaries. Polygon features are used to represent these geographic boundaries. The GIS data is in svy21 projected coordinates systems.

The code chunk below is used to import geoBoundaries-UGA-ADM2 shapefile by using st_read() of sf packages.

```{r}
# Load district level boundary GIS data

boundaries2 <- st_read(dsn = "data/rawdata/geoBoundaries-UGA-ADM2-all", 
                layer = "geoBoundaries-UGA-ADM2")
```

## 5.1.1 Updating CRS Information

Uganda is located in southeast Africa between 1º S and 4º N latitude, and between 30º E and 35º E longitude.

The code chunk below updates the newly imported mpsz with the correct ESPG code (i.e. 32736 or 21096).

```{r}
# Transform to the correct ESPG Code

boundaries <- st_transform(boundaries2, 32736)
```

```{r}
# Verify the newly transformed boundaries

st_crs(boundaries)
```

```{r}
st_bbox(boundaries) #view extent
```

```{r}
tm_shape(boundaries) +
  tm_polygons()
```

```{r}
## Convert to multipolygon to individual polygon
boundaries_sf <- boundaries %>% 
  st_cast("POLYGON") %>% 
  mutate(area = st_area(.))
```

```{r}
## Group by the unique name and select the largest polygon by area
boundaries_cleaned <- boundaries_sf %>% 
  group_by(shapeName) %>% 
  filter(area == max(area)) %>% 
  ungroup() %>% 
  select(-area) %>% 
  select(shapeName) %>% 
  rename(
  county_name = shapeName 
  )
```

```{r}
tm_shape(boundaries_cleaned) +
  tm_polygons()
```

## 5.2 Importing the aspatial data, FinScope Uganda

The FinScope-2023_Dataset_Final is in csv file format. The codes chunk below uses read_csv() function of readr package to import FinScope-2023_Dataset_Final into R as a tibble data frame called uganda_data.

```{r}
uganda_data <- read_csv("data/rawdata/FinScope-2023_Dataset_Final.csv")
```

### 5.2.1 Variables to Consider for Financial Inclusion

Check the column names in the uganda_data to identify the right names.

```{r}
#| eval: false
colnames(uganda_data) # Displays all column names in the dataset
```

To determine factors affecting financial inclusion, consider including the following types of variables:

::: panel-tabset
## Demographic Variables

-   Age and Age Band

-   Gender

-   Education Level

-   Mobile User

## Economic Variables

-   Income Level

-   Employment Status

## Geographic Variables

-   Urban vs. rural status

## Access Variables

-   Distance to nearest bank or financial institution from Home (Commerical Nank, SACCO and Mobile Money)

## Financial Services Variables

-   Financial Advice

-   Save Money and the channel (Commerical Nank, SACCO and Mobile Money)

-   Last amount saved

-   Borrow Money and the channel (Commerical Nank, SACCO and Mobile Money)

-   Last amount borrowed

-   Last amount sent

-   Last amount received

## Other

-   Documentation for KYC (National Identification Card, Passport, Utilities and Pay Slip)

-   Self Sustaining
:::

```{r}
uganda_data_new <- uganda_data %>%
  select(-c(2:4, 6:7, 9, 11:17, 20, 23:28, 30:34, 36:37, 40:43, 45:63, 65, 67:90, 93:94, 96, 98:167, 169:230, 232:234, 236:238, 240:241, 243:342, 344:364, 366:384, 386:438, 440:444, 446, 448:674, 677:679, 681:686)) %>% 
  rename(
    age_band = c1,
    gender = c2,
    education_level = c4,
    employment_status = c5,
    mobile_user = c7_1_1,
    national_ic_doc = c8_1a,
    passport_doc = c8_1d,
    utilities_bill_doc = c8_1e,
    pay_slip_doc = c8_1j,
    self_sustaining = e1_1,
    financial_advice = e3_1,
    save_money = f2_1,
    save_money_commercial_bank = f3_1_1,
    save_money_SACCO = f3_1_4,
    save_money_mobile_money = f3_1_6,
    last_amt_saved = f6_1,
    last_amt_borrowed = g3_3,
    borrow_money_commercial_bank = g6_1_1,
    borrow_money_SACCO = g6_1_5,
    borrow_money_mobile_money = g6_1_8,
    last_amt_sent = hpp3_2,
    last_amt_received = hpp6_2,
    own_insurance = j1,
    distance_commerical_bank = k1_1_1,
    distance_SACCOS = k1_1_7,
    distance_mobile_money = k1_1_9,
    latitude = hh_gps_latitude,
    longitude = hh_gps_longitude,
    county_name = s1aq2b
  )
```

```{r}
uganda_data_new <- uganda_data_new %>%
  filter(!is.na(longitude) & longitude != "",
         !is.na(latitude) & latitude != "")
```

```{r}
head(uganda_data_new$longitude) #see the data in XCOORD column
```

```{r}
head(uganda_data_new$latitude) #see the data in YCOORD column
```

Next, summary() of base R is used to display the summary statistics of uganda_data_new tibble data frame.

```{r}
summary(uganda_data_new)
```

## 5.3 Converting aspatial data frame into a sf object

Currently, the uganda_data_new tibble data frame is aspatial. We will convert it to a sf object. The code chunk below converts uganda_data_new data frame into a simple feature data frame by using st_as_sf() of sf packages.

```{r}
uganda_data.sf <- st_as_sf(uganda_data_new,
                           coords = c("longitude", "latitude"),
                           crs=4326) %>%
  st_transform(crs=32736) %>%
  replace_na(list(
    save_money_commercial_bank = 2,
    save_money_SACCO = 2,
    save_money_mobile_money = 2,
    last_amt_saved = 9,
    last_amt_borrowed = 998,
    borrow_money_commercial_bank = 2,
    borrow_money_SACCO = 2,
    borrow_money_mobile_money = 2,
    last_amt_sent = 998,
    last_amt_received = 998
  ))
```

Notice that st_transform() of sf package is used to convert the coordinates from wgs84 (i.e. crs:4326) to Arc 1960 (i.e. crs=32736).

Next, head() is used to list the content of uganda_data.sf object.

```{r}
head(uganda_data.sf)
```
